### 2018-04-05 kaggle.Rmd
Add new files

---
title: "MSDS 6371 Kaggle Project"
author: "Team Awesome"
date: "April 5, 2018"
output: html_document
---
output: 
  html_document:
    toc: true
    toc_depth: 2
  pdf_document:
    toc: true
    toc_depth: 2
---

```{r setup, include=FALSE, echo=TRUE}
library(knitr)

# Clean the workspace, remove all existing Objects
rm(list=ls())
knitr::opts_chunk$set(echo = TRUE)

# Perform environment setup
source('code/00_Setup.R')

library(xlsx)
source('code/01_LoadData.R')
# Creates dataframe:  hr_employee_attrition
```

## R Markdown

This is an R Markdown document. Markdown is a simple formatting syntax for authoring HTML, PDF, and MS Word documents. For more details on using R Markdown see <http://rmarkdown.rstudio.com>.

When you click the **Knit** button a document will be generated that includes both content as well as the output of any embedded R code chunks within the document. You can embed an R code chunk like this:

```{r cars}
summary(cars)
```

## Including Plots

You can also embed plots, for example:

```{r pressure, echo=FALSE}
plot(pressure)
```

Note that the `echo = FALSE` parameter was added to the code chunk to prevent printing of the R code that generated the plot.