### 2018-04-05 kaggle.Rmd
commit changes

---
title: "MSDS 6371 Kaggle Project"
author: "Team Awesome"
date: "April 5, 2018"
output: html_document
---
output: 
  html_document:
    toc: true
    toc_depth: 2
  pdf_document:
    toc: true
    toc_depth: 2
---

```{r setup, include=FALSE, echo=TRUE}
library(knitr)

# Clean the workspace, remove all existing Objects
rm(list=ls())
knitr::opts_chunk$set(echo = TRUE)

# Perform environment setup
source('code/00_Setup.R')

source('code/01_LoadData.R')
# Creates dataframe:  hr_employee_attrition
```

## R Markdown

## Including Plots