### 2018-04-10 msds6371-group-project-kaggle.Rmd
Add latest info and graphs

---
title: "DDSAnalytics  - Employee Trends"
author:
- Randall Hendrickson
- Rajat Chandna
- Lokesh Maganti
- Mike Kelleher
date: "April 15, 2018"
output: 
  html_document:
    toc: true
    toc_depth: 2
    fig_caption: yes
  pdf_document:
    toc: true
    toc_depth: 2
    fig_caption: yes
---

```{r setup, include=FALSE, echo=FALSE}
library(knitr)
knitr::opts_chunk$set(fig.path = 'images/', echo = FALSE)
```
![Kaggle House Prices](images/Kaggle.png)
\pagebreak

## Introduction

This is the final project for the semester in MSDS 6371 Statistical Foundation for Data Science, and is a group project.  It is based on the [Kaggle House Prices competition](https://www.kaggle.com/c/house-prices-advanced-regression-techniques).  Fundamentally we are seeking to answer what will the sale price of a house be based on some combination of predictive attribute measures of it.  The 2 specific questions below prescribe distinct approaches to answering this question.

## Data Description

As mentioned above, the data comes from the [Kaggle House Prices competition](https://www.kaggle.com/c/house-prices-advanced-regression-techniques).  The specific data sets we use for this project are the training data set (train.csv) to build the models, and test data set (test.csv) to cross-validate our models.

### train.csv

This dataset is roughly 450k with 1460 observations of 81 variables.  To read and understand more about this specific dataset please refer to the Kaggle website at [https://www.kaggle.com/c/house-prices-advanced-regression-techniques/data](https://www.kaggle.com/c/house-prices-advanced-regression-techniques/data).

### test.csv

This dataset, is roughly 441k with 1459 observation of 80 variables.  This has one less observation: SalesPrices.  This is due to the fact it is meant to be used for cross-validation, and does not require this variable because we are meant to predict that with our models.

### Primary Variables

The following variables are fundamental to the answer for Question 1, do not contain missing values, and require no cleaning:
* SalePrice - the property's sale price in dollars
* GrLivArea - Above grade (ground) living area square feet
* Neighborhood - Physical locations within Ames city limits

## Question 1

### The Problem

Ames Century 21 real estate would like us to perform the following analysis of homes:

For the neighborhoods of: BrkSide, Edwards and NAmes, please find a predictive model for home SalePrice based on GrLivArea.  Provide the model assumptions assessment with evidence, as well as data review and outlier analysis.  Please provide estimate or estimates with confidence intervals, and a written conclusion of the relationship of GrLivArea and SalePrice within these neighborhoods.

#### Untransformed Data

When looking at plots of the raw data, there is a clear linear relationship between GrLivArea and SalePrice.  However, we can start to see there may be some issues with some of the points on Image 002, Image 003, and Image 004.  These will be explored later.

Most of the parameter estimates look good except for d3, int2 and int3.  However the VIF for all estimates are high (See Image 005), the smallest being 22.  This is another sign that outliers should be investigated.

When looking at the residual plots (Image 006) we can clearly see that Equal Variance is violated when looking athe the Residual and Studentized Residual.  Also, the QQPlot and Histogram displays data that may not be normal.

We need to perform outlier analysis and peform transformations to address equal deviation and normality.

#### Transformed Data

The following transformations were performed to attempt to address the issues found with the untransformed data: Normality and Equal Variance.  Both Log-Linear and Linear-Log did not address these issues with the data.  The final transformation Log-Log did.

##### Log-Log Transformation

When looking at the Log-Log plots (Images 007 to 010) Linearity still holds.  R^2 increased from .57 to .58, and all the VIF's decreased significantly, but some are still above 10.  The d3 and int3 variables are still not statistically significnt, but int2 is now statistically significant.

When looking at residual plots (Image 012), everything looks much better.  Residuals now have equal variance.  The QQ Plot has MUCh less curve to it.  The histogram looks more normal, but still has a left skew.  Cooks-D shows a few points might have some influence on our model.

#### Outlier Analysis

### Assumptions
* Linear Trend (Plots: Response vs Explanatory, Explanatory vs Explanatory)
* Normality (Residual: Scatter, QQ Plot, Histogram)
* Equal Standard Deviation (Residual Scatter Plots)

### Comparison of Competing Models

### Parameters
* Estimates
* Interpretation
* Confidence Intervals

### Conclusion

## Question 2

### The Problem

### Model Selection

#### Forward

##### Assumptions
* Linear Trend (Plots: Response vs Explanatory, Explanatory vs Explanatory)
* Normality (Residual: Scatter, QQ Plot, Histogram)
* Equal Standard Deviation (Residual Scatter Plots)

#### Backward

##### Assumptions
* Linear Trend (Plots: Response vs Explanatory, Explanatory vs Explanatory)
* Normality (Residual: Scatter, QQ Plot, Histogram)
* Equal Standard Deviation (Residual Scatter Plots)

#### Stepwise

##### Assumptions
* Linear Trend (Plots: Response vs Explanatory, Explanatory vs Explanatory)
* Normality (Residual: Scatter, QQ Plot, Histogram)
* Equal Standard Deviation (Residual Scatter Plots)

#### Custom

##### Assumptions
* Linear Trend (Plots: Response vs Explanatory, Explanatory vs Explanatory)
* Normality (Residual: Scatter, QQ Plot, Histogram)
* Equal Standard Deviation (Residual Scatter Plots)

### Comparison of Models
| Predictive Models | Adusted R^2 | CV Press | Kaggle Score |
|-------------------|:-----------:|:--------:|:------------:|
| Forward           | 0.1111      | 1111     | .111         |
| Backward          | 0.2222      | 2222     | .222         |
| Stepwise          | 0.3333      | 3333     | .333         |
| Custom            | 0.4444      | 4444     | .444         |

### Conclusion

## Appendix

### Question 1

#### Plots of Untransformed Data
![BrkSide](images/Untransformed-BrkSide.png)*Image 001*
![Edwards](images/Untransformed-Edwards.png)*Image 002*
![NAmes](images/Untransformed-NAmes.png)*Image 003*
![Other Neighborhoods](images/Untransformed-Others.png)*Image 004*
![MLR Parameter Estimates](images/Untransformed-MLR-Estimates.png)*Image 005*
![MLR Residual Plots](images/Untransformed-MLR-Residuals.png)*Image 006*

----

#### Plots of Log-Log Data
![BrkSide](images/Log-Log-BrkSide.png)*Image 007*
![Edwards](images/Log-Log-Edwards.png)*Image 008*
![NAmes](images/Log-Log-NAmes.png)*Image 009*
![Other Neighborhoods](images/Log-Log-Others.png)*Image 010*
![MLR Parameter Estimates](images/Log-Log-MLR-Estimates.png)*Image 011*
![MLR Residual Plots](images/Log-Log-MLR-Residuals.png)*Image 012*

----

### Question 2